---
title: '2.4 SFARI Genes'
output:
  html_document:
    code_folding: 'hide'
---

<br>

#### Load R packages

```{r load_packages, warning=FALSE, message=FALSE}
library(tidyverse) ; library(reshape2) ; library(glue) ; library(plotly)
library(RColorBrewer) ; library(viridis) ; require(gridExtra) ; library(GGally) ; library(ggpubr)
library(rstatix)
library(expss)
library(DESeq2)
library(knitr) ; library(kableExtra)
```

#### Load preprocessed dataset (code in 2.1.Preprocessing_pipeline.Rmd)

```{r load_dataset, echo=TRUE, message=FALSE, warning=FALSE}

# SFARI Genes
SFARI_genes = read_csv('./../../SFARI/Data/SFARI_genes_01-03-2020_w_ensembl_IDs.csv')

# Load Gandal dataset
load('./../Data/preprocessedData/preprocessed_data.RData')
datExpr = datExpr %>% data.frame

# Updates genes_info with SFARI information
genes_info = genes_info %>% left_join(SFARI_genes, by = 'ID') %>%
             mutate(gene.score = ifelse(is.na(`gene-score`) & Neuronal==0, 'Others', 
                                        ifelse(is.na(`gene-score`), 'Neuronal', `gene-score`))) %>%
             mutate(Group = factor(ifelse(gene.score %in% c('Neuronal','Others'), gene.score, 'SFARI'), 
                    levels = c('SFARI', 'Neuronal', 'Others')))

# SFARI palette
SFARI_colour_hue = function(r) {
  pal = c('#FF7631','#FFB100','#E8E328','#8CC83F','#62CCA6','#59B9C9','#b3b3b3','#808080','gray','#d9d9d9')[r]
}

```
<br>

------

<br>

There are `r sum(SFARI_genes$ID %in% rownames(datExpr))` SFARI Genes in the expression dataset (~`r round(100*mean(SFARI_genes$ID %in% rownames(datExpr)))`%), `r sum(!genes_info[['gene-score']] %in% c('None','Others'))` of them have a SFARI Gene Score.

Gene count by SFARI score:
```{r}
table_info = genes_info %>% apply_labels(`gene-score` = 'SFARI Gene Score', syndromic = 'Syndromic Tag',
                                          Neuronal = 'Neuronal Function', gene.score = 'Gene Score') %>%
             mutate(syndromic = as.logical(syndromic), Neuronal = as.logical(Neuronal))

cro(table_info$`gene-score`)
```
<br>

Gene count by Syndromic tag:
```{r}
cro(table_info$syndromic)
```
<br>

Neuronal annotations:

<br>

`r sum(genes_info$Neuronal)`  genes have neuronal-related annotations, `r sum(genes_info$gene.score %in% c('1','2','3') & genes_info$Neuronal)` of these, have a SFARI score

```{r}
cro(table_info$gene.score[genes_info$`gene-score` %in% as.character(c(1:3))],
    list(table_info$Neuronal[genes_info$`gene-score` %in% as.character(c(1:3))], total()))

rm(table_info)
```
<br>

-----

<br>

## 2.4.1 Analysis of all SFARI Genes together
<br>

### Mean Expression

```{r, fig.align='center'}

plot_data = data.frame('ID'=rownames(datExpr), 'MeanExpr'=rowMeans(datExpr)) %>% 
            left_join(genes_info, by='ID')

wt = plot_data %>% wilcox_test(MeanExpr~Group, p.adjust.method='BH') %>% add_x_position(x = 'group')
increase = 1
base = 15.7
pos_y_comparisons = c(base, base+increase, base)
plot_data %>% ggplot(aes(Group, MeanExpr)) + 
     geom_boxplot(outlier.colour='#cccccc', outlier.shape='o', outlier.size=3, aes(fill=Group)) + 
     stat_pvalue_manual(wt, y.position = pos_y_comparisons, tip.length = .03) +
     scale_fill_manual(values=c('#00A4F7', SFARI_colour_hue(r=c(8,7)))) + 
     xlab('') + ylab('Mean Expression') + #ggtitle('Mean Expression Comparison') +
     scale_x_discrete(labels=c('SFARI' = 'SFARI\nGenes', 'Others' = 'Other\ngenes',
                              'Neuronal' = 'Neuronal\ngenes')) +
     theme_minimal() + theme(legend.position='none')

```

### Log Fold Change

The propotion of over and underexpressed genes is very simiar for each group, so we can just focus on the magnitude of the LFC independently of the sign.

```{r, fig.align='center'}

genes_info %>% mutate(direction = ifelse(log2FoldChange>0, 'overexpressed', 'underexpressed')) %>% 
               group_by(Group, direction) %>% tally(name = 'overexpressed') %>% 
               filter(direction == 'overexpressed') %>% ungroup %>% 
               left_join(genes_info %>% group_by(Group) %>% tally(name = 'Total'), by = 'Group') %>% ungroup %>%
               mutate('underexpressed' = Total - overexpressed , 
                      'ratio' = round(overexpressed/underexpressed,2)) %>% 
               dplyr::select(Group, overexpressed, underexpressed, ratio) %>% 
               kable %>% kable_styling(full_width = F)

```
<br>

```{r, fig.align='center'}

plot_data = data.frame('ID'=rownames(datExpr), 'MeanExpr'=rowMeans(datExpr)) %>% 
            left_join(genes_info, by='ID')

wt = plot_data %>% filter(log2FoldChange>0) %>% 
     wilcox_test(MeanExpr~Group, p.adjust.method='BH') %>% add_x_position(x = 'group')
increase = 1
base = 15.7
pos_y_comparisons = c(base, base+increase, base)
p1 = plot_data %>% filter(log2FoldChange>0) %>% ggplot(aes(Group, MeanExpr)) +
     geom_boxplot(outlier.colour='#cccccc', outlier.shape='o', outlier.size=3, aes(fill=Group)) + 
     stat_pvalue_manual(wt, y.position = pos_y_comparisons, tip.length = .03) +
     scale_fill_manual(values=c('#00A4F7', SFARI_colour_hue(r=c(8,7)))) + 
     xlab('') + ylab('Mean Expression') + ggtitle('Overexpressed genes') +
     scale_x_discrete(labels=c('SFARI' = 'SFARI\nGenes', 'Others' = 'Other\ngenes',
                              'Neuronal' = 'Neuronal\ngenes')) +
     theme_minimal() + theme(legend.position='none')

wt = plot_data %>% filter(log2FoldChange<0) %>% 
     wilcox_test(MeanExpr~Group, p.adjust.method='BH') %>% add_x_position(x = 'group')
p2 = plot_data %>% filter(log2FoldChange<0) %>% ggplot(aes(Group, MeanExpr)) +
     geom_boxplot(outlier.colour='#cccccc', outlier.shape='o', outlier.size=3, aes(fill=Group)) + 
     stat_pvalue_manual(wt, y.position = pos_y_comparisons, tip.length = .03) +
     scale_fill_manual(values=c('#00A4F7', SFARI_colour_hue(r=c(8,7)))) + 
     xlab('') + ylab('Mean Expression') + ggtitle('Underexpressed genes') +
     scale_x_discrete(labels=c('SFARI' = 'SFARI\nGenes', 'Others' = 'Other\ngenes',
                              'Neuronal' = 'Neuronal\ngenes')) +
     theme_minimal() + theme(legend.position='none')

grid.arrange(p1, p2, nrow=1)

```

<br>

Both with the original LFC values as well as the shrunken ones, the difference in LFC magnitude between SFARI genes and Neuronal genes is statisically significant with a p-value lower than $10^{-4}$.

```{r fig.align='center'}

wt = genes_info %>% mutate(abs_lfc = abs(log2FoldChange), 
                           Group = factor(Group, levels = c('SFARI','Others', 'Neuronal'))) %>% 
     wilcox_test(abs_lfc~Group, p.adjust.method='BH') %>% add_x_position(x = 'group')

increase = 0.04
base = 0.45
pos_y_comparisons = c(base, base + increase, base)
p1 = genes_info %>% mutate(Group = factor(Group, levels = c('SFARI','Others', 'Neuronal'))) %>% 
     ggplot(aes(Group, abs(log2FoldChange))) + 
     geom_boxplot(outlier.colour='#cccccc', outlier.shape='o', outlier.size=3, aes(fill=Group)) + 
     stat_pvalue_manual(wt, y.position = pos_y_comparisons, tip.length = .005) + 
     scale_fill_manual(values=c('#00A4F7', SFARI_colour_hue(r=c(7,8)))) + 
     coord_cartesian(ylim= c(0, max(pos_y_comparisons))) + 
     scale_x_discrete(labels=c('SFARI' = 'SFARI\nGenes', 'Others' = 'Other\ngenes',
                              'Neuronal' = 'Neuronal\ngenes')) +
     xlab('') + ylab('Original LFC Magnitude') + theme_minimal() + theme(legend.position='none')


wt = genes_info %>% mutate(abs_lfc = abs(shrunken_log2FoldChange), 
                           Group = factor(Group, levels = c('SFARI','Others', 'Neuronal'))) %>% 
     wilcox_test(abs_lfc~Group, p.adjust.method='BH') %>% add_x_position(x = 'group')

increase = 0.03
base = 0.35
pos_y_comparisons = c(base, base + increase, base)
p2 = genes_info %>% mutate(Group = factor(Group, levels = c('SFARI','Others', 'Neuronal'))) %>% 
     ggplot(aes(Group, abs(shrunken_log2FoldChange))) + 
     geom_boxplot(outlier.colour='#cccccc', outlier.shape='o', outlier.size=3, aes(fill=Group)) + 
     stat_pvalue_manual(wt, y.position = pos_y_comparisons, tip.length = .01) + 
     scale_fill_manual(values=c('#00A4F7', SFARI_colour_hue(r=c(7,8)))) + 
     coord_cartesian(ylim= c(0, max(pos_y_comparisons))) + 
     scale_x_discrete(labels=c('SFARI' = 'SFARI\nGenes', 'Others' = 'Other\ngenes',
                              'Neuronal' = 'Neuronal\ngenes')) +
     xlab('') + ylab('Shrunken LFC Magnitude') + theme_minimal() + theme(legend.position='none')

grid.arrange(p1, p2, nrow = 1)

rm(increase, base, pos_y_comparisons, wt)
```

Full plots, without cropping out outliers

```{r fig.align='center'}

p1 = ggplotly(genes_info %>% mutate(Group = factor(Group, levels = c('SFARI','Others', 'Neuronal'))) %>% 
     ggplot(aes(Group, abs(log2FoldChange))) + 
     geom_boxplot(outlier.colour='#cccccc', outlier.shape='o', outlier.size=3, aes(fill=Group)) + 
     scale_fill_manual(values=c('#00A4F7', SFARI_colour_hue(r=c(7,8)))) + 
     scale_x_discrete(labels=c('SFARI' = 'SFARI\nGenes', 'Others' = 'Other\ngenes',
                              'Neuronal' = 'Neuronal\ngenes')) +
     xlab('') + ylab('Original LFC Magnitude') + theme_minimal() + theme(legend.position='none'))
  
p2 = ggplotly(genes_info %>% mutate(Group = factor(Group, levels = c('SFARI','Others', 'Neuronal'))) %>% 
     ggplot(aes(Group, abs(shrunken_log2FoldChange))) + 
     geom_boxplot(outlier.colour='#cccccc', outlier.shape='o', outlier.size=3, aes(fill=Group)) + 
     scale_fill_manual(values=c('#00A4F7', SFARI_colour_hue(r=c(7,8)))) + 
     scale_x_discrete(labels=c('SFARI' = 'SFARI\nGenes', 'Others' = 'Other\ngenes',
                              'Neuronal' = 'Neuronal\ngenes')) +
     xlab('') + ylab('Shrunken LFC Magnitude') + theme_minimal() + theme(legend.position='none'))
  

subplot(p1, p2, nrows = 1)

```

<br>

### Differential Expression

```{r}
genes_info %>% group_by(Group, significant) %>% tally(name = 'DE') %>% filter(significant) %>% ungroup %>%
               left_join(genes_info %>% group_by(Group) %>% tally(name = 'Total'), by = 'Group') %>% ungroup %>%
               mutate('prop_DE' = round(DE/Total,2)) %>% dplyr::select(Group, DE, prop_DE, Total) %>% 
               kable(caption = 'Proportion of DE Genes by Group') %>% kable_styling(full_width = F)
```
<br>

```{r warning=FALSE, message=FALSE, fig.width=10}

lfc_list = seq(0, 0.3, 0.01)

all_counts = data.frame('group'='All', 'n'=as.character(nrow(genes_info)))
Others_counts = data.frame('group'='Others', n=as.character(sum(genes_info$Group == 'Others')))
Neuronal_counts = data.frame('group'='Neuronal', n=as.character(sum(genes_info$Group == 'Neuronal')))
lfc_counts_all = genes_info %>% filter(Group == 'SFARI') %>% tally %>%
                 mutate('group'='SFARI', 'n'=as.character(n)) %>% 
                 dplyr::select(group, n) %>%
                 bind_rows(Neuronal_counts, Others_counts, all_counts) %>%
                 mutate('lfc'=-1) %>% dplyr::select(lfc, group, n)

for(lfc in lfc_list){
  
  # Recalculate genes_info with the new threshold (p-values change)
  DE_genes = results(dds, lfcThreshold=lfc, altHypothesis='greaterAbs') %>% data.frame %>% 
             mutate('ID' = rownames(.)) %>% 
             left_join(genes_info %>% dplyr::select(ID, Neuronal, gene.score, Group), by = 'ID') %>% 
             filter(padj<0.05 & abs(log2FoldChange)>lfc)

  
  # Calculate counts by groups
  all_counts = data.frame('group'='All', 'n'=as.character(nrow(DE_genes)))
  Others_counts = data.frame('group'='Others', n=as.character(sum(DE_genes$Group == 'Others')))
  Neuronal_counts = data.frame('group'='Neuronal', n=as.character(sum(DE_genes$Group == 'Neuronal')))
  lfc_counts = DE_genes %>% filter(Group == 'SFARI') %>% tally %>%
               mutate('group'='SFARI', 'n'=as.character(n)) %>%
               bind_rows(Neuronal_counts, Others_counts, all_counts) %>%
               mutate('lfc'=lfc) %>% dplyr::select(lfc, group, n)
  
  
  # Update lfc_counts_all
  lfc_counts_all = lfc_counts_all %>% bind_rows(lfc_counts)
}

# Add missing entries with 0s
lfc_counts_all = expand.grid('group'=unique(lfc_counts_all$group), 'lfc'=unique(lfc_counts_all$lfc)) %>% 
  left_join(lfc_counts_all, by=c('group','lfc')) %>% replace(is.na(.), 0)

# Calculate percentage of each group remaining
tot_counts = genes_info %>% filter(Group == 'SFARI') %>% tally() %>%
             mutate('group'='SFARI', 'tot'=n) %>% dplyr::select(group, tot) %>%
             bind_rows(data.frame('group'='Neuronal', 'tot'=sum(genes_info$Group == 'Neuronal')),
                       data.frame('group' = 'Others', 'tot' = sum(genes_info$Group == 'Others')),
                       data.frame('group'='All', 'tot'=nrow(genes_info)))

lfc_counts_all = lfc_counts_all %>% filter(lfc!=-1) %>% #, group!='Others') %>% 
                 left_join(tot_counts, by='group') %>% mutate('perc'=round(100*as.numeric(n)/tot,2))


# Plot change of number of genes
ggplotly(lfc_counts_all %>% filter(group != 'All') %>% 
         mutate(group = factor(group, levels = c('SFARI', 'Neuronal', 'Others'))) %>%
         ggplot(aes(lfc, perc, color=group)) + geom_point(aes(id=n)) + geom_line() +
         scale_color_manual(values=c('#00A4F7', SFARI_colour_hue(r=c(8,7)))) + 
         ylab('Percentage of DE Genes') +  xlab('LFC threshold') + labs(color = 'Group') + 
         #ggtitle('Effect of filtering thresholds in SFARI Genes') + 
         theme_minimal() + theme(legend.position = 'bottom'))

rm(lfc_list, all_counts, Others_counts, Neuronal_counts, lfc_counts, lfc_counts_all, DE_genes, tot_counts)

```

<br>

-----

<br>

## 2.4.2 Analysis of SFARI Genes by SFARI Score
<br>

### Mean Expression

```{r fig.align='center'}

plot_data = data.frame('ID'=rownames(datExpr), 'MeanExpr'=rowMeans(datExpr)) %>% left_join(genes_info, by='ID')

wt = plot_data %>% wilcox_test(MeanExpr~gene.score, p.adjust.method='BH') %>% add_x_position(x = 'group')
increase = 1
base = 15.5
pos_y_comparisons = c(base + c(0,1,3,5)*increase, base+c(0,2,4)*increase, base+0:1*increase, base)
plot_data %>% ggplot(aes(gene.score, MeanExpr)) + 
              geom_boxplot(outlier.colour='#cccccc', outlier.shape='o', outlier.size=3, aes(fill=gene.score)) +
              stat_pvalue_manual(wt, y.position = pos_y_comparisons, tip.length = .01) +
              scale_fill_manual(values=SFARI_colour_hue(r=c(1:3,8,7))) + 
              xlab('') + ylab('Mean Expression') + 
              scale_x_discrete(labels=c('1' = 'SFARI\nScore 1', '2' = 'SFARI\nScore 2', '3' = 'SFARI\nScore 3', 
                                        'Others' = 'Other\ngenes', 'Neuronal' = 'Neuronal\ngenes')) +
              theme_minimal() + theme(legend.position='none')

```


### Log Fold Change

```{r, fig.align='center'}

genes_info %>% mutate(direction = ifelse(log2FoldChange>0, 'overexpressed', 'underexpressed')) %>% 
               group_by(gene.score, direction) %>% tally(name = 'overexpressed') %>% 
               filter(direction == 'overexpressed') %>% ungroup %>% 
               left_join(genes_info %>% group_by(gene.score) %>% tally(name = 'Total'), by = 'gene.score') %>% ungroup %>%
               mutate('underexpressed' = Total - overexpressed , 
                      'ratio' = round(overexpressed/underexpressed,2)) %>% 
               dplyr::select(gene.score, overexpressed, underexpressed, ratio) %>% 
               kable %>% kable_styling(full_width = F)

```
<br>

```{r, fig.align='center'}

plot_data = data.frame('ID'=rownames(datExpr), 'MeanExpr'=rowMeans(datExpr)) %>% 
            left_join(genes_info, by='ID')

wt = plot_data %>% filter(log2FoldChange>0) %>% 
     wilcox_test(MeanExpr~gene.score, p.adjust.method='BH') %>% add_x_position(x = 'group')
increase = 1
base = 15.5
pos_y_comparisons = c(base + c(0,1,3,5)*increase, base+c(0,2,4)*increase, base+0:1*increase, base)
p1 = plot_data %>% filter(log2FoldChange>0) %>% ggplot(aes(gene.score, MeanExpr)) +
     geom_boxplot(outlier.colour='#cccccc', outlier.shape='o', outlier.size=3, aes(fill=gene.score)) + 
     stat_pvalue_manual(wt, y.position = pos_y_comparisons, tip.length = .03) +
     scale_fill_manual(values=SFARI_colour_hue(r=c(1:3,8,7))) + 
     xlab('') + ylab('Mean Expression') + ggtitle('Overexpressed genes') +
     scale_x_discrete(labels=c('1' = 'SFARI\nScore 1', '2' = 'SFARI\nScore 2', '3' = 'SFARI\nScore 3', 
                                        'Others' = 'Other\ngenes', 'Neuronal' = 'Neuronal\ngenes')) +
     theme_minimal() + theme(legend.position='none')

wt = plot_data %>% filter(log2FoldChange<0) %>% 
     wilcox_test(MeanExpr~gene.score, p.adjust.method='BH') %>% add_x_position(x = 'group')
p2 = plot_data %>% filter(log2FoldChange<0) %>% ggplot(aes(gene.score, MeanExpr)) +
     geom_boxplot(outlier.colour='#cccccc', outlier.shape='o', outlier.size=3, aes(fill=gene.score)) + 
     stat_pvalue_manual(wt, y.position = pos_y_comparisons, tip.length = .03) +
      scale_fill_manual(values=SFARI_colour_hue(r=c(1:3,8,7))) + 
     xlab('') + ylab('Mean Expression') + ggtitle('Underexpressed genes') +
     scale_x_discrete(labels=c('1' = 'SFARI\nScore 1', '2' = 'SFARI\nScore 2', '3' = 'SFARI\nScore 3', 
                                        'Others' = 'Other\ngenes', 'Neuronal' = 'Neuronal\ngenes')) +
     theme_minimal() + theme(legend.position='none')

grid.arrange(p1, p2, nrow=1)

```

<br>

Both with the original LFC values as well as the shrunken ones, the difference in LFC magnitude between SFARI genes and Neuronal genes is statisically significant with a p-value lower than $10^{-4}$.

```{r fig.align='center'}

wt = genes_info %>% mutate(abs_lfc = abs(log2FoldChange), 
                           Group = factor(gene.score, levels = c('1','2','3','Others','Neuronal'))) %>% 
     wilcox_test(abs_lfc~Group, p.adjust.method='BH') %>% add_x_position(x = 'group')

increase = 0.04
base = 0.45
pos_y_comparisons = c(base + c(0,1,3,5)*increase, base+c(0,2,4)*increase, base+0:1*increase, base)
p1 = genes_info %>% mutate(gene.score = factor(gene.score, levels = c('1','2','3','Others','Neuronal'))) %>%
     ggplot(aes(gene.score, abs(log2FoldChange))) + 
     geom_boxplot(outlier.colour='#cccccc', outlier.shape='o', outlier.size=3, aes(fill=gene.score)) + 
     stat_pvalue_manual(wt, y.position = pos_y_comparisons, tip.length = .005) + 
     scale_fill_manual(values=SFARI_colour_hue(r=c(1:3,7,8))) + 
     coord_cartesian(ylim= c(0, max(pos_y_comparisons))) + 
     scale_x_discrete(labels=c('1' = 'SFARI\nScore 1', '2' = 'SFARI\nScore 2', '3' = 'SFARI\nScore 3', 
                               'Neuronal' = 'Neuronal\ngenes','Others' = 'Other\ngenes')) +
     xlab('') + ylab('Original LFC Magnitude') + theme_minimal() + theme(legend.position='none')


wt = genes_info %>% mutate(abs_lfc = abs(shrunken_log2FoldChange), 
                           Group = factor(gene.score, levels = c('1','2','3','Others', 'Neuronal'))) %>% 
     wilcox_test(abs_lfc~Group, p.adjust.method='BH') %>% add_x_position(x = 'group')

increase = 0.03
base = 0.35
pos_y_comparisons = c(base + c(0,1,3,5)*increase, base+c(0,2,4)*increase, base+0:1*increase, base)
p2 = genes_info %>% mutate(gene.score = factor(gene.score, levels = c('1','2','3','Others','Neuronal'))) %>%
     ggplot(aes(gene.score, abs(shrunken_log2FoldChange))) + 
     geom_boxplot(outlier.colour='#cccccc', outlier.shape='o', outlier.size=3, aes(fill=gene.score)) + 
     stat_pvalue_manual(wt, y.position = pos_y_comparisons, tip.length = .01) + 
     scale_fill_manual(values=SFARI_colour_hue(r=c(1:3,7,8))) + 
     coord_cartesian(ylim= c(0, max(pos_y_comparisons))) + 
     scale_x_discrete(labels=c('1' = 'SFARI\nScore 1', '2' = 'SFARI\nScore 2', '3' = 'SFARI\nScore 3', 
                               'Neuronal' = 'Neuronal\ngenes', 'Others' = 'Other\ngenes')) +
     xlab('') + ylab('Shrunken LFC Magnitude') + theme_minimal() + theme(legend.position='none')

grid.arrange(p1, p2, nrow = 1)

rm(increase, base, pos_y_comparisons, wt)
```
<br>

Full plots, without cropping out outliers

```{r fig.align='center'}

p1 = genes_info %>% mutate(gene.score = factor(gene.score, levels = c('1','2','3','Others','Neuronal'))) %>%
     ggplot(aes(gene.score, abs(log2FoldChange))) + 
     geom_boxplot(outlier.colour='#cccccc', outlier.shape='o', outlier.size=3, aes(fill=gene.score)) + 
     scale_fill_manual(values=SFARI_colour_hue(r=c(1:3,7,8))) + 
     scale_x_discrete(labels=c('1' = 'SFARI\nScore 1', '2' = 'SFARI\nScore 2', '3' = 'SFARI\nScore 3', 
                               'Neuronal' = 'Neuronal\ngenes','Others' = 'Other\ngenes')) +
     xlab('') + ylab('Original LFC Magnitude') + theme_minimal() + theme(legend.position='none')

p2 = genes_info %>% mutate(gene.score = factor(gene.score, levels = c('1','2','3','Others','Neuronal'))) %>%
     ggplot(aes(gene.score, abs(shrunken_log2FoldChange))) + 
     geom_boxplot(outlier.colour='#cccccc', outlier.shape='o', outlier.size=3, aes(fill=gene.score)) + 
     scale_fill_manual(values=SFARI_colour_hue(r=c(1:3,7,8))) + 
     scale_x_discrete(labels=c('1' = 'SFARI\nScore 1', '2' = 'SFARI\nScore 2', '3' = 'SFARI\nScore 3', 
                               'Neuronal' = 'Neuronal\ngenes', 'Others' = 'Other\ngenes')) +
     xlab('') + ylab('Shrunken LFC Magnitude') + theme_minimal() + theme(legend.position='none')

subplot(ggplotly(p1), ggplotly(p2), nrows=1)

rm(increase, base, pos_y_comparisons, wt)
```
<br>

### Differential Expression

<br>

```{r lfc_by_SFARI_score_plot, fig.width=10, warning=FALSE, message=FALSE}

lfc_list = seq(0, 0.3, 0.01)

all_counts = data.frame('group'='All', 'n'=as.character(nrow(genes_info)))
Others_counts = data.frame('group'='Others', n=as.character(sum(genes_info$gene.score == 'Others')))
Neuronal_counts = data.frame('group'='Neuronal', n=as.character(sum(genes_info$gene.score == 'Neuronal')))
lfc_counts_all = genes_info %>% group_by(`gene-score`) %>% filter(`gene-score` != 'Others') %>% tally %>%
                 mutate('group'=as.factor(`gene-score`), 'n'=as.character(n)) %>%
                 dplyr::select(group, n) %>%
                 bind_rows(Neuronal_counts, Others_counts, all_counts) %>%
                 mutate('lfc'=-1) %>%  dplyr::select(lfc, group, n)

for(lfc in lfc_list){
  
  # Recalculate genes_info with the new threshold (p-values change)
  DE_genes = results(dds, lfcThreshold=lfc, altHypothesis='greaterAbs') %>% data.frame %>%
             mutate('ID'=rownames(.)) %>% 
             left_join(genes_info %>% dplyr::select(ID, Neuronal, gene.score), by='ID') %>% 
             filter(padj<0.05 & abs(log2FoldChange)>lfc)

  
  # Calculate counts by groups
  all_counts = data.frame('group'='All', 'n'=as.character(nrow(DE_genes)))
  Others_counts = data.frame('group'='Others', n=as.character(sum(DE_genes$gene.score == 'Others')))
  Neuronal_counts = data.frame('group'='Neuronal', n=as.character(sum(DE_genes$gene.score == 'Neuronal')))
  lfc_counts = DE_genes %>% group_by(gene.score) %>% tally %>%
               mutate('group'=gene.score, 'n'=as.character(n)) %>%
               bind_rows(Neuronal_counts, all_counts) %>%
               mutate('lfc'=lfc) %>% dplyr::select(lfc, group, n)
  
  
  # Update lfc_counts_all
  lfc_counts_all = lfc_counts_all %>% bind_rows(lfc_counts)
}

# Add missing entries with 0s
lfc_counts_all = expand.grid('group'=unique(lfc_counts_all$group), 'lfc'=unique(lfc_counts_all$lfc)) %>% 
  left_join(lfc_counts_all, by=c('group','lfc')) %>% replace(is.na(.), 0)

# Calculate percentage of each group remaining
tot_counts = genes_info %>% group_by(`gene-score`) %>% tally() %>% filter(`gene-score`!='Others') %>%
             mutate('group'=factor(`gene-score`), 'tot'=n) %>% dplyr::select(group, tot) %>%
             bind_rows(data.frame('group'='Neuronal', 'tot'=sum(genes_info$gene.score == 'Neuronal')),
                       data.frame('group'='Others', 'tot'=sum(genes_info$gene.score == 'Others')),
                       data.frame('group'='All', 'tot'=nrow(genes_info)))

lfc_counts_all = lfc_counts_all %>% filter(lfc!=-1) %>% #, group!='Others') %>% 
                 left_join(tot_counts, by='group') %>% mutate('perc'=round(100*as.numeric(n)/tot,2))


# Plot change of number of genes
ggplotly(lfc_counts_all %>% filter(group != 'All') %>% 
         mutate(group = ifelse(group %in% c('1','2','3'), paste0('Score ',group), group)) %>% 
         mutate(group = factor(group, levels = c('Score 1', 'Score 2', 'Score 3', 'Neuronal', 'Others'))) %>%
         ggplot(aes(lfc, perc, color=group)) + 
         geom_point(aes(id=n)) + geom_line() + scale_color_manual(values=SFARI_colour_hue(r=c(1:3,8,7))) + 
         ylab('Percentage of DE Genes') +  xlab('LFC threshold') +  labs(color = 'Group') +
         #ggtitle('Effect of filtering thresholds by SFARI score') + 
         #guides(color=guide_legend(nrow=2,byrow=TRUE) + # for breaking the legend into two rows
         theme_minimal() + theme(legend.position = 'bottom'))

rm(lfc_list, all_counts, Neuronal_counts, lfc_counts_all, lfc, lfc_counts, lfc_counts_all, tot_counts,
   lfc_counts_all, Others_counts)
```

---

#### Session info

```{r}
sessionInfo()
```
<br><br>
